---
title: "Multilevel Model for Biology"
author: "Brendi Ang"
date: "17/10/2021"
output: 
  bookdown::html_document2:
    theme: flatly
    toc: true
    number_sections: false
---

```{r bio-setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r, include=FALSE}
library(tidyverse)
library(broom)
library(broom.mixed)
library(colorspace)
library(kableExtra)
library(lme4)
library(patchwork)

ggplot2::theme_set(theme_bw())

select <- dplyr::select
```

```{r, include=FALSE, warning=FALSE}
# Read-in data
mathsci_all <- read_csv(here::here("data/mathsci_all.csv"),
                        col_types = cols(doe_centre_code = col_factor()))

# Convert data types
mathsci_all <- mathsci_all %>% 
  mutate(across(.cols = c(qcaa_subject_id, 
                          subject_type,
                          qcaa_school_id,
                          sector,
                          school_postcode
                          ),
                ~ as.factor(.x))) 
```

# Biology

```{r}
bio <- mathsci_all %>% 
  filter(subject_name == "Biology")

bio_long <- bio %>% 
  pivot_longer(cols = year_11_enrolments:year_12_enrolments,
               names_to = "unit",
               values_to = "enrolments") 
```

## Exploring the dataset with basic linear model for each school

```{r bio-eda, fig.cap = "Basic linear model for 20 randomly selected schools to provide an at-a-glance visualisation of enrolment trends within schools for Biology subject"}
# Fit a linear model for 20 random sampled schools
set.seed(546)
bio_long %>% 
  filter(qcaa_school_id %in% sample(qcaa_school_id, size = 20),
         enrolments > 0) %>% 
  ggplot(aes(x = completion_year,
             y = enrolments)) +
  geom_point() +
  stat_smooth(method = "lm",
              colour = "orange") +
  facet_wrap(~ qcaa_school_id) +
  scale_x_continuous(breaks = seq(1992, 2022, by = 2),
                     labels = seq(1992, 2022, by = 2)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Completion year",
       y = "Enrolments",
       title = "Linear fit for 20 randomly selected schools")
```

As described in [step 1](#step-one), a basic linear model was plotted for each school to provide insights of the enrolment trends for each school. Figure \@ref(fig:bio-eda) demonstrates Biology may be introduced in schools in later years (*e.g.* school 235) and schools may have removed the subject and in some cases, school 1 and school 645 did not exist after 1997. School 570 showed a halt in the subject from 2005 to 2014. Schools can also vary greatly in enrolment sizes, for instance, schools 129, 132 and 570 had less than 50 enrolments for every cohort while some schools (*e.g.* school 235) have relatively larger cohort size. Some schools also showed a general decrease in enrolments across the years while some schools such at school 5 and 470 showed a significant increase in enrolments over the years. 

## Getting the data ready for modelling

### Removing graduating cohort 2019

All zero enrolments in a given year will be removed for modelling. As aforementioned, most of the zero enrolments in year 11 (refer to Figure \@ref(fig:year-11-enrolments-all)) were attributed to the 2007 prep year cohort while zero enrolments in year 12 relates to the first year in which a school introduces the subject. Other zero enrolments mostly relates to smaller schools with little to no enrolments in the subject for a given year.

### Linearise relationship using log transformation

```{r bio-log-scale, fig.cap = "Effects of log transformation for response variable (enrolments) in Biology subject"}
p1 <- ggplot(bio_long, aes(enrolments)) +
      geom_histogram(binwidth = 15,
                     fill = "white",
                     colour = "black") +
      geom_density(aes(y = 15 * ..count..),
                   fill = "steelblue",
                   alpha = 0.4) +
    labs(title = "Before log transformation",
         x = "Enrolments",
         y = "Density")

p2 <- ggplot(bio_long, aes(log(enrolments))) +
      geom_histogram(binwidth = 0.3,
                     fill = "white",
                     colour = "black") +
      geom_density(aes(y= 0.3 * ..count..),
                   fill = "steelblue",
                   alpha = 0.4) +
  labs(title = "After log transformation",
       x = "Log Enrolments",
       y = "Density")

p1 / p2
```

A log transformation was used on the response variable (`enrolments`) to allow model to be estimated by the multilevel model, which assumes normality in the error terms. As shown in Figure \@ref(fig:bio-log-scale).

```{r}
# ----- Cleaning script

bio <- bio %>% 
  # Remove graduating cohort 2019 
  filter(completion_year != 2019) %>% 
  # Remove schools that offer subject in one of Year 11 or Year 12 only
  filter(year_12_enrolments != 0,
         year_11_enrolments != 0)

bio_long <- bio_long %>% 
  # Remove graduating cohort 2019 
  filter(completion_year != 2019) %>% 
  # Remove schools that offer subject in one of Year 11 or Year 12 only
  filter(enrolments != 0) %>% 
  # Recentre year
  mutate(year92 = completion_year - min(completion_year),
         .before = completion_year) %>% 
  # Log transformation of response variable
  mutate(log_enrolments = log(enrolments))
```

## Unconditional means model

```{r}
# ----- Fit possible null model

# Two-level: Within schools
model0.0 <- lmer(log_enrolments ~ 1 + (1 | qcaa_school_id),
                 data = bio_long)

# Three-level: Schools nested within postcodes
model0.1 <- lmer(log_enrolments ~ 1 + (1 | school_postcode/qcaa_school_id),
                 data = bio_long)

# Three-level Schools nested within districts
model0.2 <- lmer(log_enrolments ~ 1 + (1 | qcaa_district/qcaa_school_id),
                 data = bio_long)
```

```{r bio-init-models}
# Obtain AIC for each null model
model0_AIC <- AIC(model0.0, model0.1, model0.2) 

# Change row names 
rownames(model0_AIC) <- c("Model0.0: Within schools",
                          "Model0.1: Schools nested within postcodes",
                          "Model0.2: Schools nested within districts")

# Show AIC for each candidate model
model0_AIC %>% 
  arrange(AIC) %>% 
  kable(caption = "AIC values for all candidate models for Biology") %>% 
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hovered", "striped"))
```

Referring back to [step 3](#step-three), three candidate models are fitted, with the AIC shown in Table \@ref(tab:bio-init-models). `Model0.2`, corresponding to having schools nested within districts is the best model, with optimised (lowest) AIC and will be used in the subsequent analysis.

### Intraclass correlation ($ICC$)

```{r, eval=FALSE, echo=TRUE}
summary(model0.2)
```

```{r}
# --- Model summary output

# Extract variance and correlation component
vc <- VarCorr(model0.2)

cat("Random effects:", "\n")

# Print variance and standard deviation of random effects
print(vc, comp = c("Variance", "Std.Dev."))

cat("\n","Fixed effects:", "\n")
# Extract fixed effects coefficient
coef(summary(model0.2))

# Print grouping structure
cat("\n",
    "Number of schools (level-two group) =",
    summary(model0.2)$ngrps[1], "\n", 
    "Number of district (level-three group) =",
    summary(model0.2)$ngrps[2])
```


In a three-level multilevel model, two intraclass correlations can be obtained using the model summary output above:

The **level-two ICC** relates to the correlation between school $i$ from a certain district $k$ in time $t$ and in time $t^*$:

$$\text{Level-two ICC} = \frac{\tau^{2}_{00}}{\tau^{2}_{00} + \phi^{2}_{00} + \sigma^2} = \frac{0.6153}{(0.6153 + 0.1113 + 0.2204)} = 0.6497$$

This can be conceptualised as the correlation between enrolments of two random draws from the same school at two different years. In other words, 64.97% of the total variability is attributable to the differences between schools from the same district rather than changes over time within schools.

The **level-three ICC** refers to the correlation between different schools $i$ and $i^*$ from a specific school $j$. 

$$\text{Level-three ICC} = \frac{\phi^2_{00}}{\tau^2_{00} + \phi^2_{00} + \sigma^2} = \frac{0.1113}{(0.6153 + 0.1113 + 0.1113)} = 0.1175$$

Similarly, it can be inferred that the correlation between enrolments of two randomly selected schools from different districts are 11.75%, where the total variability can be attributed to the difference between districts.

## Unconditional growth model

```{r}
# Unconditional growth model
model1.0 <- lmer(log_enrolments ~ year92 + 
                (year92 | qcaa_district:qcaa_school_id) + 
                (year92 | qcaa_district),
                REML = TRUE,
                control = lmerControl(optimizer = "bobyqa"),
                data = bio_long)
```

```{r, eval=FALSE, echo=TRUE}
summary(model1.0)
```

```{r}
# --- Model summary output

# Extract variance and correlation component
vc <- VarCorr(model1.0)

# Print variance and standard deviation of random effects
print(vc, comp = c("Variance", "Std.Dev."))

# Extract fixed effects coefficient
coef(summary(model1.0))

# Print grouping structure
cat(" Number of Level Two groups = ",
    summary(model1.0)$ngrps[1], "\n", 
    "Number of Level Three groups = ",
    summary(model1.0)$ngrps[2])
```

The unconditional growth model adds the systematic changes over time, the model specification can be found in [step 4](#step-four). This allows for assessing within-school variability which can be attributed to the linear changes over time. Based on the model output:

- $\pi_{0ij}$ = 2.6826: Initial status for school $i$ in district $j$ (*i.e.* expected log enrolments when time = 0)
- $\pi_{1ij}$ = 0.0259: Growth rate for school $i$ in district $j$
- $\epsilon_{tij}$ = 0.1503: Variance in within-school residuals after accounting for linear growth overtime

Biology was first introduced in 1992, and schools are expected to have 14.62 ($e^{2.68260}$), on average. Furthermore, enrolments were expected to increase by 2.628\% ($e^{0.02594} - 1) \times 100$) every year. The estimated within-school variance decrease by 24.49\% (0.2204 to 0.1503), implying that 24.49\% of the within-school variability can be explained by the linear growth over time.

## Testing fixed effects 

```{r}
model4.0 <- lmer(log_enrolments ~ year92*sector*unit +
                (year92 | qcaa_district:qcaa_school_id) + 
                (year92 | qcaa_district),
                REML = FALSE,
                control = lmerControl(optimizer = "bobyqa"),
                data = bio_long)

# ----- Specify all models with different fixed effects

# --- Remove three-way interaction

model4.1 <- update(model4.0, . ~ . - sector:unit:year92)

# --- Remove two two-way interaction

model4.2 <- update(model4.1, . ~. - sector:year92 - unit:year92)

model4.3 <- update(model4.1, . ~. - unit:sector - sector:year92)

model4.4 <- update(model4.1, . ~. - unit:sector - unit:year92)

# --- Remove one two-way interaction

# Remove sector:unit interaction
model4.5 <- update(model4.1, . ~ . - sector:unit)

# Remove sector:year92 interaction
model4.6 <- update(model4.1, . ~ . - sector:year92)

# Remove unit:year92 interaction
model4.7 <- update(model4.1, . ~ . - unit:year92)

# --- No interactions

# Include 3 fixed effects
model4.8 <- update(model4.1, . ~. - unit:year92 - sector:year92 - unit:year92)

# Remove unit fixed effects
model4.9 <- update(model4.8, . ~. - unit)

# Remove sector fixed effects
model4.10 <- update(model4.8, . ~. - sector)
```

```{r}
# ANOVA test
model_anova <- anova(
  model4.0,
  model4.1,
  model4.2,
  model4.3,
  model4.4,
  model4.5,
  model4.6,
  model4.7,
  model4.8,
  model4.9,
  model4.10
)

model_anova %>% 
  as_tibble() %>% 
  select(npar, AIC, BIC, logLik) %>% 
  mutate(model = rownames(model_anova),
         .before = npar) %>% 
  arrange(AIC)
```

As detailed in [step 6](#step-six), level-two predictors (`sector` and `unit`) are added to the model. The largest possible model will be fitted, before removing each fixed effect one by one whilst recording the AIC for each model. `model4.0` corresponds to the largest model while `model4.10` is the smallest possible model. The model with the optimal (lowest) AIC is `model4.1`, and will be used in subsequent sections.

## Parametric bootstrap to test random effects

```{r}
bootstrapAnova <- function(mA, m0, B = 1000){
  oneBootstrap <- function(m0, mA){
    
    # Regenerate new set of response with the null model
    d <- drop(simulate(m0))
    
    # Fit both null and full model to the new data
    m2 <- refit(mA, newresp=d)
    m1 <- refit(m0, newresp=d)
    
    # Return chisq test statistic
    return(anova(m2,m1)$Chisq[2])
  }  
  
  # Conduct test 1,000 times and
  # Store chisq test statistic for each iteration
  nulldist <- replicate(B, oneBootstrap(m0, mA))
  
  # chisq statistic based on the proposed and null model
  ret <- anova(mA, m0)
  
  # Change p-value to bootstrap p-value
  # Proportion of times simulated chisq statistic > actual chisq test statistic
  ret$"Pr(>Chisq)"[2] <- mean(ret$Chisq[2] < nulldist)
  
  # Change name from Pr(>Chisq) to Pr_boot(>Chisq)
  names(ret)[8] <- "Pr_boot(>Chisq)"
  
  # Change heading of model output
  attr(ret, "heading") <- c(attr(ret, "heading")[1], 
    paste("Parametric bootstrap with", B,"samples."),
    attr(ret, "heading")[-1])
  
  attr(ret, "nulldist") <- nulldist
  return(ret)
}
```

```{r}
# Best model based on AIC

# Proposed model
balt <- lmer(log_enrolments ~ year92 + sector + unit + year92:sector + year92:unit + sector:unit +
               (year92 | qcaa_district:qcaa_school_id) +  
               (year92 | qcaa_district),
             REML = FALSE,
             data = bio_long)

# Null model
bnull <- lmer(log_enrolments ~ year92 + sector + unit + year92:sector + year92:unit + sector:unit +
               (year92 | qcaa_district:qcaa_school_id) +  
               (1 | qcaa_district),
              REML = FALSE,
              data = bio_long)

actual <- 2*(logLik(balt) - logLik(bnull))
```

```{r, eval=FALSE, warning=FALSE, message=FALSE}
bootstrapLRT <- bootstrapAnova(mA = balt,
                               m0 = bnull, 
                               B = 1000)

save(bootstrapLRT, file = here::here("data/bootstrapLRT/bio.rda"))
```

```{r bio-bootstrapLRT-tab}
load(here::here("data/bootstrapLRT/bio.rda"))

as_tibble(bootstrapLRT) %>% 
  kable(caption = "Parametric Bootstrap to compare larger and smaller, nested model") %>% 
  kable_styling(full_width = FALSE)
```


```{r bio-bootstrapLRT-plot, fig.cap = "Histogram of likelihood ratio test statistic, with a red vertical line indicating the likelihood ratio test statistic for the actual model"}
nullLRT = attr(bootstrapLRT, "nulldist")
x = seq(0, max(nullLRT), length = 100)
y = dchisq(x, 2)

nullLRT.1 <- as.data.frame(cbind(nullLRT = nullLRT, x = x, y = y))

ggplot(nullLRT.1) +
  geom_histogram(
    aes(x = nullLRT, y = ..density..),
    bins = 25,
    color = "black",
    fill = "white"
  ) + 
  geom_vline(xintercept = actual, 
             size = 1,
             colour = "red") +
  geom_line(aes(x = x, y = y)) +
  labs(y = "Density",
       x = "Likelihood Ratio Test Statistics from Null Distribution") 
```

The parametric bootstrap is used to approximate the likelihood ratio test statistic to produce a more accurate p-value by simulating data under the null hypothesis (detailed explanation can be found in [step 7](#step-seven). Figure \@ref(fig:bio-bootstrapLRT-plot) displays the likelihood ratio test statistic from the null distribution, with the red line representing the likelihood ratio test statistic using the actual data. The p-value of 14.5\% (Table \@ref(tab:bio-bootstrapLRT-tab)) indicates the proportion of times in which the bootstrap test statistic is greater than the observed test statistic. The large estimated $p$-value is 0.145 < 0.05 fails to reject the null hypothesis at the 5% level, indicating that the smaller model (without random slope at level three) is preferred.


## Confidence interval

```{r}
# Re-Fit best model with restricted maximum likelihood estimates
model_f <- lmer(log_enrolments ~ year92 + sector + unit + year92:sector + year92:unit + sector:unit +
                  (year92 | qcaa_district:qcaa_school_id) +  
                  (1 | qcaa_district),
             REML = TRUE,
             control = lmerControl(optimizer = "bobyqa"),
             data = bio_long)
```

```{r, eval=FALSE}
confint_f <- confint(model_f,
        method = "boot",
        oldNames = FALSE)

save(confint_f, file = here::here("data/confint/bio.rda"))
```

```{r bio-confint}
load(here::here("data/confint/bio.rda"))

as_tibble(confint_f) %>% 
  mutate(var = rownames(confint_f),
         .before = `2.5 %`) %>% 
  kable(caption = "95% confidence intervals for fixed and random effects in the final model") %>% 
  kable_styling(full_width = FALSE)
```

The parametric bootstrap is utilised to construct confidence intervals (as detailed in [step 8](#step-eight)). If the confidence intervals for the random effects does not include 0, it provides statistical evidence that the p-value is less than 0.5. In other words, it suggests that the random effects and the correlation between the random effects are significant at the 5% level. 

The 95% confidence interval is shown above (Table \@ref(tab:bio-confint)), and the random effects all exclude 0, further reiterating that they are statistically significant at the 5% level. Some fixed effects such as `unityear_12_enrolments` were insignificant, suggesting that there were no differences between unit 11 and unit 12 units.

## Interpreting the final model

### Composite model


- Level one (measurement variable)
$$Y_{tij} = \pi_{0ij} + \pi_{1ij}year92_{tij} + \epsilon_{tij}$$

- Level two (schools within districts) 
  $$\pi_{0ij} = \beta_{00j} + \beta_{01j}sector_{ij} + \beta_{02j}unit_{ij} + \beta_{03j}sector_{ij}unit_{ij} + u_{0ij} \\ 
    \pi_{1ij} = \beta_{10j} + \beta_{11j}sector_{ij} + u_{1ij}$$

- Level three (districts)
  $$\beta_{00j} = \gamma_{000} + r_{00j} \\
    \beta_{01j} = \gamma_{010} + r_{01j} \\
    \beta_{02j} = \gamma_{020} + r_{02j} \\
    \beta_{03j} = \gamma_{030} + r_{03j} \\
    \beta_{10j} = \gamma_{100} \\
    \beta_{11j} = \gamma_{110} $$

Therefore, the composite model can be written as

$$\begin{aligned} Y_{tij} =\ &\pi_{0ij} + \pi_{1ij}year92_{tij} + \epsilon_{tij} \\ =\ &(\beta_{00j} + \beta_{01j}sector_{ij} + \beta_{02j}unit_{ij} + \beta_{03j}sector_{ij}unit_{ij} + u_{0ij}) + 
           (\beta_{10j} + \beta_{11j}sector_{ij} + u_{1ij})year92_{tij} + \epsilon_{tij} \\ =\ & \left[(\gamma_{000} + r_{00j}) + (\gamma_{010} + r_{01j})sector_{ij} + (\gamma_{020} + r_{02j})unit_{ij} + (\gamma_{030} + r_{03j})sector_{ij}unit_{ij} + u_{0ij} \right] + \\ & \left[\gamma_{100} + \gamma_{110}sector_{ij} + u_{1ij} \right]year92_{tij} + \epsilon_{tij} \\
=\ & \left[\gamma_{000} + \gamma_{010}sector_{ij} + \gamma_{020}unit_{ij} + \gamma_{020}sector_{ij}unit_{ij} + \gamma_{100}year92_{tij} + \gamma_{110}year92_{tij}sector_{ij} \right] + \\ 
& \left[ r_{00j} + r_{01j}sector_{ij} + r_{02j}unit_{ij} + r_{03j}sector_{ij}unit_{ij} + u_{0ij} + u_{1ij} \epsilon_{tij}
 \right] \end{aligned}$$

### Fixed effects

```{r}
# Extract fixed effects for final model
fixef_f <- fixef(model_f)

# \gamma_{000} + \gamma_{010}sector_{ij} + \gamma_{020}unit_{ij} + \gamma_{020}sector_{ij}unit_{ij} + 
# \gamma_{100}year92_{tij} + \gamma_{110}year92_{tij}sector_{ij}

# Catholic (unit 11)
fit_cat_unit11 <- fixef_f[[1]] + fixef_f[[2]] * c(0:30)

# Catholic (unit 12)
fit_cat_unit12 <- fixef_f[[1]] + fixef_f[[5]] + fixef_f[[2]] * c(0:30) 
  
# Government (unit 11)
fit_gov_unit11 <- fixef_f[[1]] + fixef_f[[3]] + fixef_f[[2]] * c(0:30) + fixef_f[[6]] * c(0:30)

# Government (unit 12)
fit_gov_unit12 <- fixef_f[[1]] + fixef_f[[3]] + fixef_f[[5]] + fixef_f[[9]] + fixef_f[[2]] * c(0:30) + fixef_f[[6]] * c(0:30)
  
# Independent (unit 11)
fit_ind_unit11 <- fixef_f[[1]] + fixef_f[[4]] + fixef_f[[2]] * c(0:30) + fixef_f[[7]] * c(0:30)
  
# Independent (unit 12)
fit_ind_unit12 <- fixef_f[[1]] + fixef_f[[4]] + fixef_f[[5]] + fixef_f[[9]] + fixef_f[[2]] * c(0:30) + fixef_f[[7]] * c(0:30)
```

```{r}
fit_f <-
  tibble(
    # Unit 11
    Catholic_unit11 = fit_cat_unit11,
    Government_unit11 = fit_gov_unit11,
    Independent_unit11 = fit_ind_unit11,
    # Unit 12
    Catholic_unit12 = fit_cat_unit12,
    Government_unit12 = fit_gov_unit12,
    Independent_unit12 = fit_ind_unit12,
    # Time variables
    year92 = 0:30,
    completion_year = year92 + 1992
  )

# Convert to long form for plotting
fit_f <- fit_f %>% 
  pivot_longer(cols = Catholic_unit11:Independent_unit12,
               names_to = "sector_unit",
               values_to = "fit") 

fit_f <- fit_f %>% 
  # Separate sector and unit into two variables
  separate(sector_unit, 
           into = c("sector", "unit"), 
           sep = "_",
           remove = FALSE) # Keep old variable for grouping in plot
```

```{r, eval=FALSE, echo=TRUE}
summary(model_f)
```

```{r}
# --- Model summary output

# Extract variance and correlation component
vc <- VarCorr(model_f)

# Print variance and standard deviation of random effects
print(vc, comp = c("Variance", "Std.Dev."))

# Extract fixed effects coefficient
coef(summary(model_f))

# Print grouping structure
cat(" Number of Level Two groups = ",
    summary(model_f)$ngrps[1], "\n", 
    "Number of Level Three groups = ",
    summary(model_f)$ngrps[2])
```

Using the model output above (see [step 9](##step-nine) for detailed explanation on fixed effects), the estimated increase in mean enrolments for government schools is 1.0229\% ($e^{0.0264 - 0.0162} - 1) \times  100$), which is 1.6402\% ($(e^{0.01626900} - 1) \times 100$) less than that of catholic schools. On the other hand, the mean enrolments for independent schools are estimated to increase by 5.9238\% ($e^{0.0264 - 0.0162} - 1) \times  100$) each year, which is 3.1592\% more than catholic schools.

```{r bio-fixef, fig.cap = "Fixed effects of the final model for Biology subject"}
fit_f %>% 
  ggplot(aes(x = completion_year,
             y = fit,
             group = sector_unit,
             colour = sector)) +
  geom_line(aes(linetype = unit),
            size = 1,
            alpha = 0.7) +
  scale_colour_discrete_qualitative() +
  labs(title = "Fixed effects of final model",
       x = "Completion Year",
       y = "Predicted log enrolments") +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90)) +
  scale_x_continuous(breaks = seq(1992, 2022, by = 2),
                     labels = seq(1992, 2022, by = 2)) 
```

The fixed effects can be better visualised in Figure \@ref(fig:bio-fixef), independent schools appears to have the highest average increase in enrolments per year. It appears that after 2022, enrolments in independent schools are predicted to be higher than government schools, on average. Unit 11 enrolments only appears to be marginally smaller than unit 12 enrolments for government and independent schools, but appears to be the same for catholic schools. Government schools starts of (in 1992) with the highest enrolments on average, but it is matched with a slow increase in enrolments over the years.

### Random effects

```{r}
# Random effects for schools within districts
ranef_school_f <- as_tibble(ranef(model_f)) %>%
  filter(!grp %in% unique(mathsci_all$qcaa_district)) %>% 
  separate(
    col = grp,
    into = c("qcaa_district", "qcaa_school_id"),
    sep = ":",
    remove = FALSE
  ) %>% 
  # Rename intercept and slope names
  mutate(term = if_else(term == "(Intercept)",
                        true = "Random Intercept",
                        false = "Random Slope")) 

# Random effects for between districts
ranef_district_f <- as_tibble(ranef(model_f)) %>% 
  filter(grp %in% unique(mathsci_all$qcaa_district)) %>% 
  group_by(term) %>% 
  rename(qcaa_district = grp) %>% 
  # Rename intercept and slope names
  mutate(term =  as.character(term),
         term = factor(if_else(
           term == "(Intercept)",
           true = "Random Intercept",
           false = term
         )))
```

```{r bio-school-ranef, fig.cap = "Random effects for schools"}
# Plot random effect for schools nested within districts
ranef_school_f %>% 
  group_by(term) %>% 
  # Reorder group based on by conditional means for plotting
  mutate(qcaa_school_id = fct_reorder(qcaa_school_id, condval)) %>% 
  ggplot() +
  geom_point(aes(x = condval,
                 y = qcaa_school_id),
             colour = "grey30") +
  geom_vline(xintercept = 0,
             size = 1,
             colour = "red") +
  facet_wrap(~ term,
             scales = "free_x")
```

Figure \@ref(fig:bio-school-ranef) displays the random effects for a given school. It is apparent that the random intercepts and slopes are negatively correlated, where a large intercept is associated with a smaller random slope. This indicates that a larger school is associated with a smaller increase (decrease) in enrolments over the years while smaller schools are predicted to have larger increase in enrolments over the years.

```{r bio-district-ranef, fig.cap = "Random intercept for districts"}
# Plot random effects for between districts
ranef_district_f %>% 
# Reorder group based on by conditional means
  mutate(qcaa_district = fct_reorder(qcaa_district, condval)) %>% 
  ggplot() + 
  geom_point(aes(x = condval,
                 y = qcaa_district)) +
  geom_vline(xintercept = 0,
             colour = "red",
             size = 1) 
```

As the random slopes are removed, all districts are predicted to have the same increase in enrolments over the years; And as was discussed previously, this was a reasonable assumption or an otherwise perfect correlation with random slope and intercept will be fitted. Figure \@ref(fig:bio-district-ranef) demonstrates that schools in Brisbane Central has the largest enrolments while Mackay have the lowest enrolments in Biology subject, on average.

### Predictions

```{r bio-pred, fig.cap = "Model predictions for 20 randomly selected schools"}
# Predictions for 20 randomly selected schools for year 11 enrolments
set.seed(136)
bio_long %>% 
  filter(unit == "year_11_enrolments") %>% 
  filter(qcaa_school_id %in% sample(qcaa_school_id, 20)) %>% 
  mutate(pred = exp(predict(model_f, newdata = .))) %>% 
  ggplot() +
  geom_line(aes(x = completion_year,
                y = exp(log_enrolments),
                group = qcaa_school_id)) +
  geom_line(aes(x = completion_year,
                y = pred,
                group = qcaa_school_id),
            colour = "orange") +
  facet_wrap(~ qcaa_school_id)
```

Figure \@ref(fig:bio-pred) above shows the predictions for 20 randomly selected schools.













