---
title: "Multilevel Model for Biology"
author: "Brendi Ang"
date: "17/10/2021"
output: 
  bookdown::html_document2:
    theme: flatly
    toc: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(tidyverse)
library(broom)
library(broom.mixed)
library(colorspace)
library(kableExtra)
library(lme4)
library(patchwork)

ggplot2::theme_set(theme_bw())

select <- dplyr::select
```

```{r, include=FALSE, warning=FALSE}
# Read-in data
mathsci_all <- read_csv(here::here("data/mathsci_all.csv"),
                        col_types = cols(doe_centre_code = col_factor()))

# Convert data types
mathsci_all <- mathsci_all %>% 
  mutate(across(.cols = c(qcaa_subject_id, 
                          subject_type,
                          qcaa_school_id,
                          sector,
                          school_postcode
                          ),
                ~ as.factor(.x))) 
```

# The dataset

```{r}
bio <- mathsci_all %>% 
  filter(subject_name == "Biology")

bio_long <- chem %>% 
  pivot_longer(cols = year_11_enrolments:year_12_enrolments,
               names_to = "unit",
               values_to = "enrolments") 
```

# Getting the data ready for modelling

## Removing graduating cohort 2019

```{r}
bio_long %>% 
  ggplot() +
  geom_line(aes(x = completion_year,
                y = enrolments,
                group = qcaa_school_id),
            alpha = 0.3,
            colour = "grey30") +
  facet_wrap(~ unit) +
  scale_x_continuous(breaks = seq(1992, 2022, by = 2),
                     labels = seq(1992, 2022, by = 2)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Enrolments excluding 2019")
```

## Linearise relationship using log transformation

```{r}
p1 <- bio_long %>%
  ggplot(aes(x = enrolments)) +
  geom_histogram() +
  labs(title = "After log transformation")

p2 <- bio_long %>%
  ggplot(aes(x = log(enrolments))) +
  geom_histogram() +
  labs(title = "After log transformation")

p1 / p2

# Time plot after log transformation
# 20 randomly selected schools
set.seed(9254)
chem_long %>% 
  filter(unit == "year_11_enrolments",
         completion_year != 2019,
         qcaa_school_id %in% sample(qcaa_school_id, 20)) %>% 
  ggplot(aes(x = completion_year,
             y = log(enrolments),
             group = qcaa_school_id)) +
  geom_line() +
  facet_wrap(~ qcaa_school_id)
```


## Removing completion year that offer chemistry subject for year 11 students but not for year 12 (and vice versa)

```{r}
p1 <- bio %>% 
  ggplot() +
  geom_point(aes(x = year_11_enrolments,
                 y = year_12_enrolments),
             alpha = 0.3) +
  geom_abline(slope = 1,
              colour = "red") +
  labs(title = "All observations") +
  theme(title = element_text(size = 8))

p2 <- bio %>% 
  filter(year_12_enrolments != 0,
         year_11_enrolments != 0) %>% 
  ggplot() +
  geom_point(aes(x = year_11_enrolments,
                 y = year_12_enrolments),
             alpha = 0.3) +
  geom_abline(slope = 1,
              colour = "red") +
  labs(title = "Removed observations with enrolments in Year 11 but not in Year 12 (and vice versa)") +
  theme(title = element_text(size = 8))
        
p1 / p2
```

- Red line shows year 11 enrolments = year 12 enrolments
- Some schools offered Specialist mathematics subjects for Year 11, but not for Year 12 (and vice versa)
  - These zero enrolment numbers will be removed from the analysis


```{r}
bio <- bio %>% 
  # Remove graduating cohort 2019 
  filter(completion_year != 2019) %>% 
  # Remove schools that offer subject in one of Year 11 or Year 12 only
  filter(year_12_enrolments != 0,
         year_11_enrolments != 0)

bio_long <- bio_long %>% 
  # Remove graduating cohort 2019 
  filter(completion_year != 2019) %>% 
  # Remove schools that offer subject in one of Year 11 or Year 12 only
  filter(enrolments != 0)
```

```{r}
# Rescale `completion_year` variable
bio_long <- bio_long %>%
  # Recentre year
  mutate(year92 = completion_year - min(completion_year),
         .before = completion_year) %>% 
  # Log transformation of response variable
  mutate(log_enrolments = log(enrolments))
```

# Unconditional means model

```{r}
# ----- Fit possible null model

# Two-level: Within schools
model0.0 <- lmer(log_enrolments ~ 1 + (1 | qcaa_school_id),
                 data = bio_long)

# Three-level: Schools nested within postcodes
model0.1 <- lmer(log_enrolments ~ 1 + (1 | school_postcode/qcaa_school_id),
                 data = bio_long)

# Three-level Schools nested within districts
model0.2 <- lmer(log_enrolments ~ 1 + (1 | qcaa_district/qcaa_school_id),
                 data = bio_long)
```

```{r}
# Obtain AIC for each null model
model0_AIC <- AIC(model0.0, model0.1, model0.2) 

# Change row names 
rownames(model0_AIC) <- c("Model0.0: Within schools",
                          "Model0.1: Schools nested within postcodes",
                          "Model0.2: Schools nested within districts")

model0_AIC %>% 
  arrange(AIC) %>% 
  kable() %>% 
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hovered", "striped"))
```

- Based on the AIC values, the nested (three-level) model performs better than the two-level multilevel model
  - Therefore, we will pick `model0.2`, which corresponds to the model where schools are nested within districts 

## Intraclass correlation ($ICC$)

```{r}
summary(model0.2)
```


The **level-two ICC** is the correlation between a school $i$ from a certain district $k$ in time $t$ and in time $t^*$:

$$ICC(school) = \frac{\tau^{2}_{00}}{\tau^{2}_{00} + \phi^{2}_{00} + \sigma^2} = \frac{0.5718}{(0.5718 + 0.1044 + 0.2032)} = 0.6502$$

The **level-three ICC** refers to the correlation between different schools $i$ and $i^*$ from a specific school $j$. 

$$ICC(school) = \frac{\phi^2_{00}}{\tau^2_{00} + \phi^2_{00} + \sigma^2} = \frac{0.1044}{(0.5718 + 0.1044 + 0.2032)} = 0.1187$$

# Unconditional growth model

- Adding systematic changes over time

```{r}
# Unconditional growth model
model1.0 <- lmer(log_enrolments ~ year92 + 
                (year92 | qcaa_district:qcaa_school_id) + 
                (year92 | qcaa_district),
                REML = TRUE,
                control = lmerControl(optimizer = "bobyqa"),
                data = bio_long)
```

- Note: no boundary constraint

# Adding predictors to the model

## Adding `sector`

```{r}
model2.0 <- lmer(log_enrolments ~ year92 + sector + sector*year92 +
                (year92 | qcaa_district:qcaa_school_id) + 
                (year92 | qcaa_district),
                REML = TRUE,
                control = lmerControl(optimizer = "bobyqa"),
                data = bio_long)
```

### Visualising fixed effects

```{r}
fixef(model2.0)
```

## Adding `sector` and `unit`

```{r}
model3.0 <- lmer(log_enrolments ~ year92 + sector + sector*year92 + unit + unit:year92 +
                (year92 | qcaa_district:qcaa_school_id) + 
                (year92 | qcaa_district),
                REML = TRUE,
                control = lmerControl(optimizer = "bobyqa"),
                data = bio_long)
```

### Visualising fixed effects



# Testing fixed effects 

```{r}
model4.0 <- lmer(log_enrolments ~ year92*sector*unit +
                (year92 | qcaa_district:qcaa_school_id) + 
                (year92 | qcaa_district),
                REML = FALSE,
                control = lmerControl(optimizer = "bobyqa"),
                data = chem_long)

# ----- Specify all models with different fixed effects

# --- Remove three-way interaction

model4.1 <- update(model4.0, . ~ . - sector:unit:year92)

# --- Remove two two-way interaction

model4.2 <- update(model4.1, . ~. - sector:year92 - unit:year92)

model4.3 <- update(model4.1, . ~. - unit:sector - sector:year92)

model4.4 <- update(model4.1, . ~. - unit:sector - unit:year92)

# --- Remove one two-way interaction

# Remove sector:unit interaction
model4.5 <- update(model4.1, . ~ . - sector:unit)

# Remove sector:year92 interaction
model4.6 <- update(model4.1, . ~ . - sector:year92)

# Remove unit:year92 interaction
model4.7 <- update(model4.1, . ~ . - unit:year92)

# --- No interactions

# Include 3 fixed effects
model4.8 <- update(model4.1, . ~. - unit:year92 - sector:year92 - unit:year92)

# Remove unit fixed effects
model4.9 <- update(model4.8, . ~. - unit)

# Remove sector fixed effects
model4.10 <- update(model4.8, . ~. - sector)
```

```{r}
# ANOVA test
model_anova <- anova(
  model4.0,
  model4.1,
  model4.2,
  model4.3,
  model4.4,
  model4.5,
  model4.6,
  model4.7,
  model4.8,
  model4.9,
  model4.10
)

model_anova %>% 
  as_tibble() %>% 
  select(npar, AIC, BIC, logLik) %>% 
  mutate(model = rownames(model_anova),
         .before = npar) %>% 
  arrange(AIC)
```
```{r}
summary(model4.7)
```

# Parametric bootstrap to test random effects

```{r}
bootstrapAnova <- function(mA, m0, B = 1000){
  oneBootstrap <- function(m0, mA){
    
    # Regenerate new set of response with the null model
    d <- drop(simulate(m0))
    
    # Fit both null and full model to the new data
    m2 <- refit(mA, newresp=d)
    m1 <- refit(m0, newresp=d)
    
    # Return chisq test statistic
    return(anova(m2,m1)$Chisq[2])
  }  
  
  # Conduct test 1,000 times and
  # Store chisq test statistic for each iteration
  nulldist <- replicate(B, oneBootstrap(m0, mA))
  
  # chisq statistic based on the proposed and null model
  ret <- anova(mA, m0)
  
  # Change p-value to bootstrap p-value
  # Proportion of times simulated chisq statistic > actual chisq test statistic
  ret$"Pr(>Chisq)"[2] <- mean(ret$Chisq[2] < nulldist)
  
  # Change name from Pr(>Chisq) to Pr_boot(>Chisq)
  names(ret)[8] <- "Pr_boot(>Chisq)"
  
  # Change heading of model output
  attr(ret, "heading") <- c(attr(ret, "heading")[1], 
    paste("Parametric bootstrap with", B,"samples."),
    attr(ret, "heading")[-1])
  
  attr(ret, "nulldist") <- nulldist
  return(ret)
}
```

```{r}
# Best model based on AIC and BIC

# Proposed model
balt <- lmer(log_enrolments ~ year92 + sector + year92:sector + sector:unit + unit + 
                (year92 | qcaa_district:qcaa_school_id) +  
                (year92 | qcaa_district),
             REML = FALSE,
             control = lmerControl(optimizer = "bobyqa"),
             data = bio_long)

# Null model
bnull <- lmer(log_enrolments ~ year92 + sector + year92:sector + sector:unit + unit + 
                (year92 | qcaa_district:qcaa_school_id) +  
                (1 | qcaa_district),
              REML = FALSE,
              control = lmerControl(optimizer = "bobyqa"),
              data = bio_long)

actual <- 2*(logLik(balt) - logLik(bnull))
```

```{r}
# ANOVA suggests smaller model better, but check parametric bootstrap
anova(balt, bnull)
```

```{r, eval=FALSE, warning=FALSE, message=FALSE}
bootstrapLRT <- bootstrapAnova(mA = balt,
                               m0 = bnull, 
                               B = 1000)

save(bootstrapLRT, file = here::here("data/bootstrapLRT/bio-model5.0-vs-model5.1.rda"))
```

```{r}
load(here::here("data/bootstrapLRT/bio-model5.0-vs-model5.1.rda"))

nullLRT = attr(bootstrapLRT, "nulldist")
x = seq(0, max(nullLRT), length = 100)
y = dchisq(x, 2)

nullLRT.1 <- as.data.frame(cbind(nullLRT = nullLRT, x = x, y = y))

ggplot(nullLRT.1) +
  geom_histogram(
    aes(x = nullLRT, y = ..density..),
    bins = 25,
    color = "black",
    fill = "white"
  ) + 
  # geom_vline(xintercept = 1000, size = 1) +
  geom_line(aes(x = x, y = y)) +
  # xlim(c(-.5,10)) +
  labs(y = "Density",
       x = "Likelihood Ratio Test Statistics from Null Distribution") 
```






