---
title: "Multilevel Model for Psychology"
author: "Brendi Ang"
date: "17/10/2021"
output: 
  bookdown::html_document2:
    theme: flatly
    toc: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(tidyverse)
library(broom)
library(broom.mixed)
library(colorspace)
library(kableExtra)
library(lme4)
library(patchwork)

ggplot2::theme_set(theme_bw())

select <- dplyr::select
```

```{r, include=FALSE, warning=FALSE}
# Read-in data
mathsci_all <- read_csv(here::here("data/mathsci_all.csv"),
                        col_types = cols(doe_centre_code = col_factor()))

# Convert data types
mathsci_all <- mathsci_all %>% 
  mutate(across(.cols = c(qcaa_subject_id, 
                          subject_type,
                          qcaa_school_id,
                          sector,
                          school_postcode
                          ),
                ~ as.factor(.x))) 
```

```{r}
psych <- mathsci_all %>% 
  filter(subject_name == "Psychology")

psych_long <- psych %>% 
  pivot_longer(cols = year_11_enrolments:year_12_enrolments,
               names_to = "unit",
               values_to = "enrolments") 
```

```{r}
mathsci_all %>% 
  ggplot() +
  geom_line(aes(x = completion_year,
                y = year_11_enrolments,
                group = qcaa_school_id)) +
  facet_wrap(~ subject_name)
```

# Getting the data ready for modelling

## Linearise response variable using log transformation

```{r, message=FALSE, warning=FALSE}
p1 <- psych_long %>%
  ggplot(aes(x = enrolments)) +
  geom_histogram() +
  labs(title = "Before log transformation") 

p2 <- psych_long %>%
  ggplot(aes(x = log(enrolments))) +
  geom_histogram() +
  labs(title = "After log transformation") +
  xlim(0,6)

p1 / p2
```

```{r}
# Rescale `completion_year` variable
psych_long <- psych_long %>%
  # Remove 3 observations with 0 enrolments
  filter(enrolments != 0) %>% 
  # Recentre year
  mutate(year20 = completion_year - min(completion_year),
         .before = completion_year) %>% 
  # Log transformation of response variable
  mutate(log_enrolments = log(enrolments))
```

# Unconditional means model

```{r}
# ----- Fit possible null model

# Two-level: Within schools
model0.0 <- lmer(log_enrolments ~ 1 + (1 | qcaa_school_id),
                 data = psych_long)

# Three-level: Schools nested within postcodes
model0.1 <- lmer(log_enrolments ~ 1 + (1 | school_postcode/qcaa_school_id),
                 data = psych_long)

# Three-level Schools nested within districts
model0.2 <- lmer(log_enrolments ~ 1 + (1 | qcaa_district/qcaa_school_id),
                 data = psych_long)
```

```{r}
# Obtain AIC for each null model
model0_AIC <- AIC(model0.0, model0.1, model0.2) 

# Change row names 
rownames(model0_AIC) <- c("Model0.0: Within schools",
                          "Model0.1: Schools nested within postcodes",
                          "Model0.2: Schools nested within districts")

model0_AIC %>% 
  arrange(AIC) %>% 
  kable() %>% 
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hovered", "striped"))
```

- Model for schools nested ran into a boundary issue for the unconditional means model, and it attributed to the lowest AIC score.
- Based on the AIC, the best model corresponds to the model with schools nested within districts.

## Intraclass correlation ($ICC$)

```{r}
summary(model0.2)
```

The **level-two ICC** is the correlation between a school $i$ from a certain district $j$ in time $t$ and in time $t^*$:

$$ICC(school) = \frac{\tau^{2}_{00}}{\tau^{2}_{00} + \phi^{2}_{00} + \sigma^2} = \frac{0.48216}{(0.48216 + 0.09015 + 0.10038)} = 0.7168$$

The **level-three ICC** refers to the correlation between different schools $i$ and $i^*$ from a specific district $j$. 

$$ICC(school) = \frac{\phi^2_{00}}{\tau^2_{00} + \phi^2_{00} + \sigma^2} = \frac{0.09015}{(0.48216 + 0.09015 + 0.10038)} = 0.1340$$

# Unconditional Growth model

```{r}
# Unconditional growth model
model1.0 <- lmer(log_enrolments ~ year20 + 
                (year20 | qcaa_district:qcaa_school_id) + 
                (year20 | qcaa_district),
                REML = TRUE,
                data = psych_long)
```

### Visualising fixed effects

```{r}
fixef(model2.0)
```


## Adding `sector` and `unit`

```{r}
model3.0 <- lmer(log_enrolments ~ year20 + sector + sector:year20 + unit + unit:year20 +
                (year20 | qcaa_district:qcaa_school_id) + 
                (year20 | qcaa_district),
                REML = TRUE,
                data = psych_long)
```

### Visualising fixed effects

```{r}
fixef(model3.0)
```

# Testing fixed effects

```{r}
model4.0 <- lmer(log_enrolments ~ year20*sector*unit +
                (year20 | qcaa_district:qcaa_school_id) + 
                (year20 | qcaa_district),
                REML = FALSE,
                control = lmerControl(optimizer = "bobyqa"),
                data = psych_long)

# ----- Specify all models with different fixed effects

# --- Remove three-way interaction

model4.1 <- update(model4.0, . ~ . - sector:unit:year20)

# --- Remove two two-way interaction

model4.2 <- update(model4.1, . ~. - sector:year20 - unit:year20)

model4.3 <- update(model4.1, . ~. - unit:sector - sector:year20)

model4.4 <- update(model4.1, . ~. - unit:sector - unit:year20)

# --- Remove one two-way interaction

# Remove sector:unit interaction
model4.5 <- update(model4.1, . ~ . - sector:unit)

# Remove sector:year20 interaction
model4.6 <- update(model4.1, . ~ . - sector:year20)

# Remove unit:year20 interaction
model4.7 <- update(model4.1, . ~ . - unit:year20)

# --- No interactions

# Include 3 fixed effects
model4.8 <- update(model4.1, . ~. - unit:year20 - sector:year20 - unit:year20)

# Remove unit fixed effects
model4.9 <- update(model4.8, . ~. - unit)

# Remove sector fixed effects
model4.10 <- update(model4.8, . ~. - sector)
```


```{r}
# ANOVA test
model_anova <- anova(
  model4.0,
  model4.1,
  model4.2,
  model4.3,
  model4.4,
  model4.5,
  model4.6,
  model4.7,
  model4.8,
  model4.9,
  model4.10
)


model_anova %>% 
  as_tibble() %>% 
  select(npar, AIC, BIC, logLik) %>% 
  mutate(model = rownames(model_anova),
         .before = npar) %>% 
  arrange(AIC)
```

- Based on the AIC values, model4.6 is the best.

# Parametric bootstrap to test random effects

```{r}
bootstrapAnova <- function(mA, m0, B = 1000){
  oneBootstrap <- function(m0, mA){
    
    # Regenerate new set of response with the null model
    d <- drop(simulate(m0))
    
    # Fit both null and full model to the new data
    m2 <- refit(mA, newresp=d)
    m1 <- refit(m0, newresp=d)
    
    # Return chisq test statistic
    return(anova(m2,m1)$Chisq[2])
  }  
  
  # Conduct test 1,000 times and
  # Store chisq test statistic for each iteration
  nulldist <- replicate(B, oneBootstrap(m0, mA))
  
  # chisq statistic based on the proposed and null model
  ret <- anova(mA, m0)
  
  # Change p-value to bootstrap p-value
  # Proportion of times simulated chisq statistic > actual chisq test statistic
  ret$"Pr(>Chisq)"[2] <- mean(ret$Chisq[2] < nulldist)
  
  # Change name from Pr(>Chisq) to Pr_boot(>Chisq)
  names(ret)[8] <- "Pr_boot(>Chisq)"
  
  # Change heading of model output
  attr(ret, "heading") <- c(attr(ret, "heading")[1], 
    paste("Parametric bootstrap with", B,"samples."),
    attr(ret, "heading")[-1])
  
  attr(ret, "nulldist") <- nulldist
  return(ret)
}
```


```{r}
# Best model based on AIC and BIC

# Proposed model
balt <- lmer(log_enrolments ~ year20 + sector + unit + + year20:unit + sector:unit +
               (year20 | qcaa_district:qcaa_school_id) +  
               (year20 | qcaa_district),
             REML = FALSE,
             control = lmerControl(optimizer = "bobyqa"),
             data = psych_long)

# Null model
bnull <- lmer(log_enrolments ~ year20 + sector + unit + + year20:unit + sector:unit +
               (year20 | qcaa_district:qcaa_school_id) +  
               (1 | qcaa_district),
              REML = FALSE,
              control = lmerControl(optimizer = "bobyqa"),
              data = psych_long)

actual <- 2*(logLik(balt) - logLik(bnull))
```

```{r, eval=FALSE, warning=FALSE, message=FALSE}
bootstrapLRT <- bootstrapAnova(mA = balt,
                               m0 = bnull, 
                               B = 1000)

save(bootstrapLRT, file = here::here("data/bootstrapLRT/psych-model5.0-vs-model5.1.rda"))
```


```{r}
# Load bootstrap likelihood ratio test output
load(here::here("data/bootstrapLRT/psych-model5.0-vs-model5.1.rda"))

as_tibble(bootstrapLRT) %>% 
  kable() %>% 
  kable_styling(full_width = FALSE)
```

- Overwhelming statistical evidence ($\chi^2$ = 0.0000, $p$-value = 0.981) supporting the smaller model.

```{r}
nullLRT = attr(bootstrapLRT, "nulldist")
x = seq(0, max(nullLRT), length = 100)
y = dchisq(x, 2)

nullLRT.1 <- as.data.frame(cbind(nullLRT = nullLRT, x = x, y = y))

ggplot(nullLRT.1) +
  geom_histogram(
    aes(x = nullLRT, y = ..density..),
    bins = 25,
    color = "black",
    fill = "white"
  ) + 
  geom_vline(xintercept = actual, 
             size = 1,
             colour = "red") +
  geom_line(aes(x = x, y = y)) +
  # xlim(c(-.5,10)) +
  labs(y = "Density",
       x = "Likelihood Ratio Test Statistics from Null Distribution") 
```

# Confidence interval

```{r}
# Fit best model with restricted maximum likelihood estimates
model_f <- lmer(log_enrolments ~ year20 + sector + unit + + year20:unit + sector:unit +
               (year20 | qcaa_district:qcaa_school_id) +  
               (1 | qcaa_district),
              REML = TRUE,
              control = lmerControl(optimizer = "bobyqa"),
              data = psych_long)
```

```{r, eval=FALSE}
confint(model_f,
        method = "boot",
        oldNames = FALSE)
```

# Interpreting final model


## Composite model 


- Level one (measurement variable)
$$Y_{tij} = \pi_{0ij} + \pi_{1ij}year20_{tij} + \epsilon_{tij}$$

- Level two (schools within districts) 
  $$\pi_{0ij} = \beta_{00j} + \beta_{01j}sector_{ij} + \beta_{02j}unit_{ij} + \beta_{03j}sector_{ij}unit_{ij} + u_{0ij} \\ 
    \pi_{1ij} = \beta_{10j} + \beta_{11j}unit{ij} + u_{1ij}$$

- Level three (districts)
  $$\beta_{00j} = \gamma_{000} + r_{00j} \\
    \beta_{01j} = \gamma_{010} + r_{01j} \\
    \beta_{02j} = \gamma_{020} + r_{02j} \\
    \beta_{03j} = \gamma_{030} + r_{03j} \\
    \beta_{10j} = \gamma_{100} \\
    \beta_{11j} = \gamma_{110} $$

Therefore, the composite model can be written as

$$\begin{aligned} 
Y_{tij} =\ &\pi_{0ij} + \pi_{1ij}year20_{tij} + \epsilon_{tij} \\

        =\ & (\beta_{00j} + \beta_{01j}sector_{ij} + \beta_{02j}unit_{ij} + \beta_{03j}sector_{ij}unit_{ij} + u_{0ij}) + \\
          & (\beta_{10j} + \beta_{11j}unit{ij})year20_{tij} + \epsilon_{tij} \\
           
        =\ & \left[\gamma_{000} + r_{00j} + (\gamma_{010} + r_{01j})sector_{ij} + (\gamma_{020} + r_{02j})unit_{ij} + (\gamma_{030} +
                  r_{03j})sector_{ij}unit_{ij} + u_{0ij} \right] + \\
           & \left[\gamma_{100} + \gamma_{110}unit_{ij} + u_{1ij} \right]year20_{tij} + \epsilon_{tij} \\
              
        =\ & \left[\gamma_{000} + \gamma_{010}sector_{ij} + \gamma_{020}unit_{ij} + \gamma_{030}sector_{ij}unit_{ij} + \gamma_{100}year20_{tij} + \gamma_{110}unit_{ij}year20_{tij} \right] \\
           & \left[r_{00j} + r_{01j}sector_{ij} + r_{02j}unit_{ij} + r_{03j}sector_{ij}unit_{ij} + u_{0ij} + u_{1ij}year20_{tij} + \epsilon_{tij} \right]
            
\end{aligned}$$ 

```{r}
# Extract fixed effects for final model
fixef_f <- fixef(model_f)

# \gamma_{000} + \gamma_{010}sector_{ij} + \gamma_{020}unit_{ij} + \gamma_{030}sector_{ij}unit_{ij} + 
# \gamma_{100}year20_{tij} + \gamma_{110}unit_{ij}year20_{tij}

# Catholic (unit 11)
fit_cat_unit11 <- fixef_f[[1]] + fixef_f[[2]] * c(0:5) 

# Catholic (unit 12)
fit_cat_unit12 <-
  fixef_f[[1]] + fixef_f[[5]] + fixef_f[[2]] * c(0:5) + fixef_f[[6]] * c(0:5)

# Government (unit 11)
fit_gov_unit11 <-
  fixef_f[[1]] + fixef_f[[3]] + fixef_f[[2]] * c(0:5)

# Government (unit 12)
fit_gov_unit12 <-
  fixef_f[[1]] + fixef_f[[3]] + fixef_f[[5]] + fixef_f[[7]] + fixef_f[[2]] * c(0:5) + fixef_f[[6]] * c(0:5)
  
# Independent (unit 11)
fit_ind_unit11 <-
  fixef_f[[1]] + fixef_f[[4]] + fixef_f[[2]] * c(0:5)
  
# Independent (unit 12)
fit_ind_unit12 <-
  fixef_f[[1]] + fixef_f[[4]] + fixef_f[[5]] + fixef_f[[8]] + fixef_f[[2]] * c(0:5) + fixef_f[[6]] * c(0:5)
```

```{r}
fit_f <-
  tibble(
    # Unit 11
    Catholic_unit11 = fit_cat_unit11,
    Government_unit11 = fit_gov_unit11,
    Independent_unit11 = fit_ind_unit11,
    # Unit 12
    Catholic_unit12 = fit_cat_unit12,
    Government_unit12 = fit_gov_unit12,
    Independent_unit12 = fit_ind_unit12,
    # Time variables
    year92 = 0:5,
    completion_year = year92 + 2020
  )

# Convert to long form for plotting
fit_f <- fit_f %>% 
  pivot_longer(cols = Catholic_unit11:Independent_unit12,
               names_to = "sector_unit",
               values_to = "fit") 

fit_f <- fit_f %>% 
  # Separate sector and unit into two variables
  separate(sector_unit, 
           into = c("sector", "unit"), 
           sep = "_",
           remove = FALSE) # Keep old variable for grouping in plot
```

```{r}
fit_f %>% 
  ggplot(aes(x = completion_year,
             y = fit,
             group = sector_unit,
             colour = sector)) +
  geom_line(aes(linetype = unit),
            size = 1,
            alpha = 0.7) +
  scale_colour_discrete_qualitative() +
  labs(title = "Fixed effects of final model",
       x = "Completion Year",
       y = "Predicted log enrolments") +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90)) +
  scale_x_continuous(breaks = seq(2020, 2025, by = 1),
                     labels = seq(2020, 2025, by = 1)) 
```

## Random effects

```{r}
# Random effects for schools within districts
ranef_school_f <- as_tibble(ranef(model_f)) %>%
  filter(!grp %in% unique(mathsci_all$qcaa_district)) %>% 
  separate(
    col = grp,
    into = c("qcaa_district", "qcaa_school_id"),
    sep = ":",
    remove = FALSE
  ) %>% 
  # Rename intercept and slope names
  mutate(term = if_else(term == "(Intercept)",
                        true = "Intercept",
                        false = "Slope")) 

# Random effects for between districts
ranef_district_f <- as_tibble(ranef(model_f)) %>% 
  filter(grp %in% unique(mathsci_all$qcaa_district)) %>% 
  group_by(term) %>% 
  rename(qcaa_district = grp) %>% 
  # Rename intercept and slope names
  mutate(term =  as.character(term),
         term = factor(if_else(
           term == "(Intercept)",
           true = "Intercept",
           false = term
         )))
```

```{r}
# Plot random effect for schools nested within districts
ranef_school_f %>% 
  group_by(term) %>% 
  # Reorder group based on by conditional means for plotting
  mutate(qcaa_school_id = fct_reorder(qcaa_school_id, condval)) %>% 
  ggplot() +
  geom_point(aes(x = condval,
                 y = qcaa_school_id)) +
  geom_vline(xintercept = 0,
             size = 1,
             colour = "red") +
  facet_wrap(~ term,
             scales = "free_x")
```

```{r}
# Plot random effects for between districts
ranef_district_f %>% 
# Reorder group based on by conditional means
  mutate(qcaa_district = fct_reorder(qcaa_district, condval)) %>% 
  ggplot() + 
  geom_point(aes(x = condval,
                 y = qcaa_district)) +
  geom_vline(xintercept = 0,
             colour = "red",
             size = 1) +
  labs(title = "Intercept")
```

## Predictions

```{r}
# Predictions for 20 randomly selected schools for year 11 enrolments
set.seed(1782)

psych_long %>% 
  filter(unit == "year_11_enrolments") %>% 
  filter(qcaa_school_id %in% sample(qcaa_school_id, 16)) %>% 
  mutate(pred = predict(model_f, newdata = .)) %>% 
  ggplot() +
  geom_line(aes(x = completion_year,
                y = log(enrolments),
                group = qcaa_school_id)) +
  geom_line(aes(x = completion_year,
                y = pred,
                group = qcaa_school_id),
            colour = "orange") +
  facet_wrap(~ qcaa_school_id) +
  theme(axis.text.x = element_text(angle = 90))
```



