---
title: "Untitled"
author: "Brendi Ang"
date: "13/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data and Motivation

```{r read-in, warning=FALSE}
# ----- Read-in data

# Enrolments for all subjects from 1992 to 2019
subj_old <- read_csv(here::here("data/qcaa_stats_subj_part_1992-2019.csv")) %>% 
  # Remove problematic entries
  filter(is.na(X13)) %>% 
  select(-c(X13, X14))
```

**Problems with dataset**
- Parts of `Subject name` are located in Subject type
  - *e.g.* subject "Film, Television & New Media" categorised as "Film" in `subject_name` and "Television and New Media" in `subject_type`
- Some `year_11_enrolments` are mistaken to be School name and vice versa
  - Due to these errors, most entries are wrongly
- Fortunately, these errors can be found easily, as they have an extra variable `X13`

```{r}
# ----- Clean variable names

subj_old <- subj_old %>% 
  janitor::clean_names() %>% 
  rename(doe_centre_code = do_e_centre_code,
         # Consistent with new qce dataset
         completion_year = cohort_year) 
```

Variables for `subj_old` dataset:
- `completion_year`: Cohort's year of completion 
- `qcaa_subject_id`: Subject ID
- `subject_name`: Subject name
- `subject_type`: Type of subject (General, Applied or Short courses)
- `doe_centre_code`: Department of Education Centre Code 
- `qcaa_school_id`: School ID
- `school_name`: School Name             
- `year_11_enrolments`: Number of students enrolled in year 11
- `year_12_enrolments`: Number of students enrolled in year 12
- `sector`: School sector (Government, Independent, Catholic)
- `school_postcode`: School postcode
- `qcaa_district`: School District

```{r}
# ----- View missing values in the dataset

# View missing values for enrolment figures
subj_old %>% 
  select(year_11_enrolments:year_12_enrolments) %>% 
  vis_miss()

# Combination of missing values
subj_old %>% 
  gg_miss_upset()
```

- Missing values are plausible, probably relates to subjects in which all enrolments are year 12 enrolments
- 20 schools have missing `doe_centre_code`

```{r}
# ----- Change data types

subj_old <- subj_old %>% 
  mutate(
    completion_year = as.factor(completion_year),
    qcaa_subject_id = as.factor(qcaa_subject_id),
    subject_type = as.factor(subject_type),
    qcaa_school_id = as.factor(qcaa_school_id),
    sector = as.factor(sector),
    school_postcode = as.factor(school_postcode),
    qcaa_district = as.factor(qcaa_district)
  ) 
```

```{r}
# ----- Impute missing values in enrolment numbers with 0

subj_old <- subj_old %>% 
  # Replace missing values in enrolment with 0
  mutate(across(.cols = year_11_enrolments:year_12_enrolments, ~ replace_na(.x, 0))) %>% 
  mutate(across(.cols = year_11_enrolments:year_12_enrolments, ~ as.numeric(.x)))
```

```{r}
math_old <- subj_old %>% 
  filter(subject_name %in% c("Mathematics A", "Mathematics B", "Mathematics C")) %>% 
  # Consistent with names of new QCE system
  mutate(subject_name = case_when(
    subject_name == "Mathematics A" ~ "General Mathematics",
    subject_name == "Mathematics B" ~ "Mathematical Methods",
    subject_name == "Mathematics C" ~ "Specialist Mathematics",
  )) %>% 
  # Create new variable to distinguish maths and science
  mutate(subject = "Mathematics")

# Enrolment for Sciences
sci_old <- subj_old %>% 
  filter(subject_name %in% c("Agricultural Science",
                             "Biology",
                             "Chemistry", 
                             "Marine Science",
                             "Physics",
                             "Agricultural Practices",
                             "Aquatic Practices",
                             "Science in Practice"
                             )) %>% 
  # Create new variable to distinguish maths and science
  mutate(subject = "Sciences")

# Combine Maths and Science subjects for the old QCE system
mathsci_old <- rbind(math_old, sci_old)

# Convert data to long form for plotting
mathsci_old_long <- mathsci_old %>% 
  pivot_longer(cols = year_11_enrolments:year_12_enrolments,
               names_to = "unit",
               values_to = "enrolments") 
```
- Two science subjects (Psychology and Earth and Environment Science) were added to the syllabus

## Exploratory Data Analysis (Old QCE system)

```{r}
# ----- 
mathsci_old_long %>% 
  filter(subject == "Mathematics") %>% 
  ggplot() +
  geom_boxplot(aes(x = subject_name, y = enrolments, fill = unit)) +
  facet_wrap(~ subject_name,
             scales = "free_x")

mathsci_old_long %>% 
  filter(subject == "Mathematics") %>% 
  ggplot(aes(x = unit, y = enrolments)) +
  geom_violin(aes(fill = unit)) +
  geom_boxplot(width = 0.3,
               outlier.alpha = 0.5) +
  facet_wrap(~ subject_name) +
  colorspace::scale_fill_discrete_sequential() +
  theme(axis.text.x = element_text(angle = 75, vjust = 0.5),
        legend.position = "none") 

mathsci_old_long %>% 
  ggplot(aes(x = completion_year, y = enrolments)) + 
  geom_boxplot() +
  facet_wrap(~ subject_name, scales = "free_x")



# ----- Year 11 and Year 12 enrolments for Mathematical Methods subject
# Too many schools to identify patterns
## Do a plotly?
ggplot() +
  geom_line(aes(group = qcaa_school_id, x = completion_year, y = enrolments),
            alpha = 0.2, data = filter(mathsci_old_long,subject_name == "Mathematical Methods")) +
  geom_line(aes(group = qcaa_school_id, x = completion_year, y = enrolments), colour = "red",
            , data = filter(mathsci_old_long, subject_name == "Mathematical Methods" &
                                              doe_centre_code == 2003)) +
  facet_wrap(~unit) +
  theme(axis.text.x = element_text(angle = 90))
```
















  