---
title: "index"
author: "Brendi Ang"
date: "10/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Set global theme for all ggplot12 objects
ggplot2::theme_set(theme_bw())
```

```{r load-library, include=FALSE}
library(tidyverse)
library(janitor)
library(naniar)
library(here)
library(readxl)
```

## Data and Motivation

```{r read-in}
# ----- Read-in data

# --- Subject enrolment numbers data

# Concatenate file names
filenames <- paste0("data/","qcaa_stats_subj_part_", c("2020", "2021", "2022"), ".xlsx")

# Read-in data
subj_2020 <- readxl::read_xlsx(here::here(filenames[1]), sheet = 1, col_types = "text")
subj_2021 <- readxl::read_xlsx(here::here(filenames[2]), sheet = 1, col_types = "text")
subj_2022 <- readxl::read_xlsx(here::here(filenames[3]), sheet = 1, col_types = "text")

# Subject enrolments based on new QCE system
subj_new <- rbind(subj_2020, subj_2021, subj_2022) %>% 
  # Remove spaces & upper-case letters in variable names
  janitor::clean_names() %>% 
  rename(doe_centre_code = do_e_centre_code)
```

- `completion_year`: Cohort's year of completion 
- `qcaa_subject_id`: Subject ID
- `subject_name`: Subject name
- `subject_type`: Type of subject (General, Applied or Short courses)
- `doe_centre_code`: Department of Education Centre Code 
- `qcaa_school_id`: School ID
- `school_name`: School Name             
- `unit_1_enrolments`: Number of students enrolled in unit 1
- `unit_2_enrolments`: Number of students enrolled in unit 2
- `unit_3_enrolments`: Number of students enrolled in unit 3      
- `unit_4_enrolments`: Number of students enrolled in unit 4
- `short_course_enrolments`: Number of students enrolled as short course enrolment
- `sector`: School sector (Government, Independent, Catholic)
- `school_postcode`: School postcode
- `qcaa_district`: School District


```{r}
# ----- View missing values in the dataset

# View missing values for dataset
subj_new %>% 
  vis_miss()

# View missing values for enrolment figures
subj_new %>% 
  select(unit_1_enrolments:unit_4_enrolments) %>% 
  vis_miss()

# Combination of missing values
subj_new %>% 
  gg_miss_upset()
```

- Nauru Secondary school, missing `doe_centre_code`, `school_postcode`, `qcaa_district`

```{r}
# ----- Change data types

subj_new <- subj_new %>% 
  mutate(
    completion_year = as.factor(completion_year),
    qcaa_subject_id = as.factor(qcaa_subject_id),
    subject_type = as.factor(subject_type),
    qcaa_school_id = as.factor(qcaa_school_id),
    sector = as.factor(sector),
    school_postcode = as.factor(school_postcode),
    qcaa_district = as.factor(qcaa_district)
  ) %>% 
  # Remove Nauru Secondary school
  filter(!is.na(qcaa_district))
```

```{r}
# ----- Impute missing values in enrolment numbers with 0

subj_new <- subj_new %>% 
  # Replace missing values in enrolment with 0
  mutate(across(.cols = unit_1_enrolments:short_course_enrolments, ~ replace_na(.x, 0))) %>% 
  mutate(across(.cols = unit_1_enrolments:short_course_enrolments, ~ as.numeric(.x)))
```

```{r}
# ----- Enrolment for mathematics and science subjects from 2020 - 2021 (Sep)

# Enrolment for Mathematics
math_new <- subj_new %>% 
  filter(subject_name %in% c("General Mathematics",
                             "Mathematical Methods",
                             "Specialist Mathematics",
                             # Applied
                             "Essential Mathematics", 
                             # Short Courses
                             "Numeracy")) %>% 
  # Create new variable to distinguish maths and science
  mutate(subject = "Mathematics")

# Enrolment for Sciences
sci_new <- subj_new %>% 
  filter(subject_name %in% c(# General
                             "Agricultural Science",
                             "Biology",
                             "Chemistry", 
                             "Earth and Environmental Science",
                             "Marine Science",
                             "Physics",
                             "Psychology",
                             # Applied
                             "Agricultural Practices",
                             "Aquatic Practices",
                             "Science in Practice"
                             )) %>% 
  # Create new variable to distinguish maths and science
  mutate(subject = "Sciences")

# Combine maths and sciences enrolments
mathsci_new <- rbind(math_new, sci_new)
```

## Exploratory Data Analysis New QCE system

### Maths

```{r}
# Total enrolments per unit
math_new_long <- math_new %>% 
  pivot_longer(cols = unit_1_enrolments:unit_4_enrolments,
               names_to = "unit",
               values_to = "enrolments") 

# Parallel coordinates chart based on each subject
math_new_long %>% 
  # Remove short course subject 
  filter(subject_name != "Numeracy") %>% 
  group_by(subject_name, unit) %>% 
  # Compute total enrolments for each subject
  summarise(total_enrolments = sum(enrolments), .groups = "drop") %>% 
  ggplot() +
  geom_line(aes(x = subject_name, y = total_enrolments, group = unit, colour = unit)) +
  geom_point(aes(x = subject_name, y = total_enrolments, colour = unit)) +
  theme(axis.text.x = element_text(angle = 90)) +
  colorspace::scale_colour_discrete_sequential()
```

### Science

```{r}
# Total enrolments per unit
sci_new_long <- sci_new %>% 
  pivot_longer(cols = unit_1_enrolments:unit_4_enrolments,
               names_to = "unit",
               values_to = "enrolments") 

# Parallel coordinates chart based on each subject
sci_new_long %>% 
  # Remove short course subject 
  filter(subject_name != "Numeracy") %>% 
  group_by(subject_name, unit) %>% 
  # Compute total enrolments for each subject
  summarise(total_enrolments = sum(enrolments), .groups = "drop") %>% 
  ggplot() +
  geom_line(aes(x = subject_name, y = total_enrolments, group = unit, colour = unit)) +
  geom_point(aes(x = subject_name, y = total_enrolments, colour = unit)) +
  theme(axis.text.x = element_text(angle = 90)) +
  colorspace::scale_colour_discrete_sequential()
```


### Maths and Science

```{r}
# Convert data to long form for plotting
mathsci_new_long <- mathsci_new %>% 
  pivot_longer(cols = unit_1_enrolments:unit_4_enrolments,
               names_to = "unit",
               values_to = "enrolments") %>% 
  # Remove short course subject 
  filter(subject_name != "Numeracy")  
```

```{r, fig.height = 8}
# ----- Total enrolments for each subject, faceted by Maths and Science

mathsci_new_long %>% 
  group_by(subject_name, unit, subject) %>% 
  # Compute total enrolments for each subject
  summarise(total_enrolments = sum(enrolments), .groups = "drop") %>%
  ggplot() +
  geom_line(aes(x = subject_name, y = total_enrolments, group = unit, colour = unit)) +
  geom_point(aes(x = subject_name, y = total_enrolments, colour = unit)) +
  facet_wrap(~ subject, scales = "free", nrow = 2) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  colorspace::scale_colour_discrete_sequential()
```


```{r}
# Distribution of enrolments for each subject by sector
library(ggridges)

# By sector
mathsci_new_long %>% 
  ggplot() +
  ggridges::geom_density_ridges(aes(x = enrolments, y = subject_name, fill = sector, point_fill = sector),
                                alpha = 0.6) +
  theme_bw() +
  # scale_fill_brewer(palette = "Dark2")
  colorspace::scale_fill_discrete_qualitative()
```

- The distribution of subjects across all schools are right skewed
  - This may be attributed to the different school sizes & thus capacity of students in a subject

```{r}
# ----- Number of enrolments per subject

mathsci_new_long %>% 
  group_by(unit, subject_name) %>% 
  summarise(total_enrolments = sum(enrolments), .groups = "drop") %>% 
  ggplot() +
  geom_line(aes(x = factor(unit), y = total_enrolments, group = subject_name)) +
  geom_point(aes(x = factor(unit), y = total_enrolments)) +
  facet_wrap(~ subject_name, scale = "free")

mathsci_new_long %>% 
  group_by(unit, subject_name) %>% 
  summarise(total_enrolments = sum(enrolments), .groups = "drop") %>% 
  ggplot() +
  geom_col(aes(x = unit, y = total_enrolments)) +
  facet_wrap(~ subject_name, scales = "free")
```

```{r}
# ----- Enrolment for each subject by sector

mathsci_new_long %>% 
  group_by(subject_name,sector, unit) %>%
  # Compute total enrolment for each unit by sector
  summarise(total_enrolments = sum(enrolments), .groups = "drop") %>% 
  ggplot() +
  geom_col(aes(x = unit, y = total_enrolments, fill = sector),
           position = "dodge") +
  facet_wrap(~ subject_name, scales = "free") +
  colorspace::scale_fill_discrete_qualitative()

mathsci_new_long %>% 
  group_by(subject_name,sector, unit) %>%
  # Compute total enrolment for each unit by sector
  summarise(total_enrolments = sum(enrolments), .groups = "drop") %>% 
  ggplot() +
  geom_line(aes(x = unit, y = total_enrolments, colour = sector, group = sector)) +
  geom_point(aes(x = unit, y = total_enrolments, colour = sector, group = sector)) +
  colorspace::scale_colour_discrete_qualitative() +
  facet_wrap(~ subject_name, scales = "free")
```

### Maths and Science (By proportion)


```{r}
# ----- Proportion of students enrolled in the subject for each cohort

# Compute size of cohort for each year
cohort_size <- mathsci_new_long %>% 
  group_by(completion_year) %>% 
  summarise(cohort_size = sum(enrolments), .groups = "drop") 

# Compute proportion of students enrolled in a subject by cohort size 
## Total enrolments in a subject / size of cohort 
mathsci_new_prop <- mathsci_new_long %>% 
  group_by(subject_name, completion_year, subject) %>% 
  summarise(total_enrolments = sum(enrolments), .groups = "drop") %>% 
  left_join(cohort_size, by = "completion_year") %>% 
  mutate(enrolment_prop = (total_enrolments/cohort_size)*100) %>% 
  select(subject_name:subject, enrolment_prop)

# Parallel Coordinate plot
ggplot(mathsci_new_prop) +
  geom_line(aes(x = subject_name, y = enrolment_prop, group = completion_year, colour = completion_year)) +
  geom_point(aes(x = subject_name, y = enrolment_prop, colour = completion_year)) +
  colorspace::scale_colour_discrete_sequential() +
  theme(axis.text.x = element_text(angle = 75, vjust = 0.5))

# Side-by-side bar plot
ggplot(mathsci_new_prop) +
  geom_col(aes(x = subject_name, y = enrolment_prop, fill = completion_year),
           position = "dodge") +
  colorspace::scale_fill_discrete_sequential() +
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
# ----- Proportion of students enrolled in the subject for each cohort and unit

mathsci_new_unit_prop <- mathsci_new_long %>%
  group_by(subject_name, completion_year, unit, subject) %>% 
  summarise(total_enrolments = sum(enrolments), .groups = "drop") %>% 
  left_join(cohort_size, by = "completion_year") %>% 
  mutate(enrolment_prop = (total_enrolments/cohort_size)*100) %>% 
  select(subject_name:subject, enrolment_prop)

# Parallel Coordinate Plot
mathsci_new_unit_prop %>% 
  # Concatenate unit and completion_year to group by cohort
  mutate(unit = paste0(unit, "_", completion_year),
         unit = factor(unit)) %>% 
  ggplot() +
  geom_line(aes(x = subject_name, y = enrolment_prop, group = unit, color = unit)) +
  colorspace::scale_colour_discrete_sequential() + 
  theme(axis.text.x = element_text(angle = 75, vjust = 0.5))

# Parallel Coordinate Plot, separating maths and science
mathsci_new_unit_prop %>% 
  select(subject_name:subject, enrolment_prop) %>% 
  # Concatenate unit and completion_year to group by cohort
  mutate(unit = paste0(unit, "_", completion_year),
         unit = factor(unit)) %>% 
  ggplot() +
  geom_line(aes(x = subject_name, y = enrolment_prop, group = unit, color = unit)) +
  facet_wrap(~ subject, scales = "free_x") +
  colorspace::scale_colour_discrete_sequential() + 
  theme(axis.text.x = element_text(angle = 75, vjust = 0.5))
```



## Problems with Dataset

```{r, warning=FALSE, eval=FALSE}
# ---- Problem with CSV file
subj_old <- read_csv(here::here("data/qcaa_stats_subj_part_1992-2019.csv")) %>% janitor::clean_names()

# ----- Problematic variables

subj_old %>% 
  filter(!is.na(x13)) %>% 
  view()
```

- Parts of `Subject name` are located in Subject type
  - *e.g.* subject "Film, Television & New Media" categorised as "Film" in `subject_name` and "Television and New Media" in `subject_type`
- Some `year_11_enrolments` are mistaken to be School name and vice versa
  - Due to these errors, most entries are wrongly
- Fortunately, these errors can be found easily, as they have an extra variable `X13`
  




