---
title: "school_outcomes"
author: "Brendi Ang"
date: "27/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# School outcomes dataset

```{r, include=FALSE}
# Read-in math & science subjects data
mathsci_all <- read_csv(here::here(data = "data/mathsci_all.csv"))
```

```{r, include=FALSE}
# 2012 to 2013
outcomes_2012 <- read_csv(here::here("data/year_12_outcomes/qsa_stats_yr12_outcomes_12.csv")) %>% 
  mutate(completion_year = 2012,
         .before = "School")

outcomes_2013 <- read_csv(here::here("data/year_12_outcomes/qsa_stats_yr12_outcomes_13.csv")) %>% 
  mutate(completion_year = 2013,
         .before = "School")
```

```{r, include=FALSE}
# 2014 to 2029
filepath_2008_2019 <- here::here(paste0("data/year_12_outcomes/qcaa_stats_yr12_outcomes_", 
                                        setdiff(seq(8,19,1), 12:13), # All numbers from 8-19 except 12 & 13 
                                        "_all_schools.csv"))

# Read-in file for each data
files <- filepath_2008_2019 %>%
  purrr::map(read_csv)

for (i in 1:length(files)) {
  
  # For first iteration
  if (i == 1) {
    outcomes_2008_2019 <-
      files[[i]] %>% mutate(completion_year = 2008, .before = "School")
  }
  
  temp <- files[[i]]
  
  # For 2009-2013
  if (i %in% 2:4){
    temp <- temp %>% mutate(completion_year = i + 2007)
    outcomes_2008_2019 <- rbind(outcomes_2008_2019, temp)
  } 
  
  # for 2010-2019
  if (i %in% 5:10){
    temp <- temp %>% mutate(completion_year = i + 2009)
    outcomes_2008_2019 <- rbind(outcomes_2008_2019, temp)
  }
}
```

```{r}
# Combine dataset from 2008-2020
school_outcomes <- rbind(outcomes_2008_2019, outcomes_2012, outcomes_2013) %>% 
  arrange(completion_year)
```

```{r}
mathsci_all <- mathsci_all %>% 
  # Remove parentheses relating to suburb
  mutate(school_name = str_remove(school_name, " \\(.*\\)")) %>% 
  # Remove suburb by extracting all text before the last "-"
  mutate(school_name = str_replace(school_name, pattern = '(.*)\\s+\\-.*', '\\1')) %>% 
  mutate(school_name = if_else(school_name == "Cloncurry State School P",
                               true = "Cloncurry State School",
                               false = school_name)) %>% 
  mutate(school_name = if_else(school_name == "Bayside Christian College Hervey Bay",
                               true = "Bayside Christian College",
                               false = school_name)) %>% 
  # Change name of school to newest name
  mutate(school_name = if_else(school_name == "Tropical North Learning Academy",
                               true = "Smithfield State High School",
                               false = school_name)) %>% 
  mutate(school_name = if_else(school_name == "St Catherine's Catholic College The Whitsundays",
                               true = "St Catherine's Catholic College",
                               false = school_name)) %>% 
  mutate(school_name = if_else(school_name == "Aboriginal & Islander Independent Community School",
                               true = "Aboriginal and Islander Independent Community School",
                               false = school_name)) %>% 
  mutate(school_name = if_else(school_name == "The SCOTS PGC College",
                               true = "Scots PGC College",
                               false = school_name)) %>% 
  mutate(school_name = if_else(str_detect(school_name, "Marist College"),
                               true = "Marist College",
                               false = school_name)) %>% 
  mutate(school_name = if_else(school_name == "The Cathedral School of St Anne and St James",
                                true = "The Cathedral School",
                                false = school_name)) %>% 
  mutate(school_name = if_else(school_name == "St Andrew's Catholic College Redlynch Valley",
                                true = "St Andrew's Catholic College",
                                false = school_name)) %>% 
  mutate(school_name = if_else(school_name == "Spinifex State College Mount Isa",
                                true = "Spinifex State College",
                                false = school_name)) %>% 
  mutate(school_name = if_else(school_name == "Cloncurry State School P-12",
                                true = "Cloncurry State School",
                                false = school_name)) %>% 
  # Capital K in MacKillop
  mutate(school_name = if_else(school_name == "Clairvaux MacKillop College",
                               true = "Clairvaux Mackillop College",
                               false = school_name)) %>% 
  # Extra " ' " in Girls
  mutate(school_name = if_else(school_name == "St Aidan's Anglican Girls' School",
                                true = "St Aidan's Anglican Girls School",
                                false = school_name))
```

```{r}
school_outcomes <- school_outcomes %>%
  # Remove schools not in enrolment data
  filter(!School %in% c("Carinity Education - Rockhampton",
                        "Carinity Education - Glendyne")) %>% 
  # Fix error for Cloncurry State school
  mutate(School = if_else(School == "Cloncurry State School P - 12",
                               true = "Cloncurry State School",
                               false = School)) %>% 
  # Fix error for Moreton Bay Boys' College
  mutate(School = if_else(School == "Moreton Bay Boys College",
                               true = "Moreton Bay Boys' College",
                               false = School)) %>% 
  mutate(School = if_else(School == "Carinity Education - Southside",
                               true = "Carinity Education",
                               false = School)) %>% 
  mutate(School = if_else(School == "Eagleby Learning College",
                               true = "Eagleby Learning Centre",
                               false = School)) %>% 
  mutate(School = if_else(School == "St Mary's Catholic College South Burnett",
                               true = "St Mary's Catholic College",
                               false = School)) %>%
  mutate(School = if_else(School == "St Catherine's Catholic College The Whitsundays",
                               true = "St Catherine's Catholic College",
                               false = School)) %>%
  mutate(School = if_else(School == "Carlisle Adventist Christian College",
                               true = "Carlisle Adventist College",
                               false = School)) %>% 
  mutate(School = if_else(School == "Agnew School",
                               true = "OneSchool Global QLD",
                               false = School)) %>%
  mutate(School = if_else(School == "Clairvaux MacKillop College",
                               true = "Clairvaux Mackillop College",
                               false = School)) %>% 
  mutate(School = if_else(School == "Emmanuel College Carrara",
                               true = "Emmanuel College",
                               false = School)) %>% 
  mutate(School = if_else(School == "St Aidan's Anglican Girls' School",
                               true = "St Aidan's Anglican Girls School",
                               false = School)) %>% 
  mutate(School = if_else(School == "St Joseph's College Gregory Terrace",
                               true = "St Joseph's College",
                               false = School)) %>% 
  mutate(School = if_else(School == "St John's College Nambour",
                               true = "St John's College",
                               false = School)) %>% 
  mutate(School = if_else(School == "St Joseph's College Toowoomba",
                               true = "St Joseph's College",
                               false = School)) %>% 
  mutate(School = if_else(School == "St Joseph's School Stanthorpe",
                               true = "St Joseph's College",
                               false = School)) %>% 
  mutate(School = if_else(School == "The Scots PGC College",
                               true = "Scots PGC College",
                               false = School)) %>%
  mutate(School = if_else(str_detect(School, "Marist College"),
                               true = "Marist College",
                               false = School)) %>%
  mutate(School = if_else(School == "The Cathedral College Rockhampton",
                               true = "The Cathedral College",
                               false = School)) %>%
  mutate(School = if_else(School == "Spinifex State College Mount Isa",
                                true = "Spinifex State College",
                                false = School)) %>% 
  # Extract all text before the last "-"
  mutate(School = str_replace(School, pattern = '(.*)\\s+\\-.*', '\\1')) %>%   
  # Remove parentheses relating to suburb
  mutate(School = str_remove(School, " \\(.*\\)")) 
```


**Schools existed before 2008 but not after 2008**
1	Acacia Ridge State High School
6	Salisbury State High School
51	Mount Carmel College
95	Newmarket State High School
96	Oxley Secondary College
97	Richlands State High School
99	Toowong College
123	Mary McConnel School
222	Frawley College
237	John Oxley College
250	Downs Secondary College
323	Slade School
332	SQIT Roma College
374	Qld International Heritage College
383	Kingaroy Christian College
553	Mount Carmel Christian Brothers College
616 Mount Isa State High School changed to Spinifex State College Mount Isa in 2003
645	Kalkadoon State High School

**Schools existing after 2020**
261 Good Samaritan College
322	St Joseph's School
538	St Teresa's College
589	St Stephen's Catholic College
607	Redwood College
611	Bellbird Park State Secondary College
623	McAuley College
694	Mary MacKillop Catholic College
698	Silver Lining School
707	Redeemer Lutheran College, Biloela
2371 Townsville Christian College
2420 Wisdom College

**Schools fixed using dashes "-"**
58	Rivermount College
341	Eagleby Learning Centre
384	Bayside Christian College
280	Carinity Education

**Schools fixed using parentheses "()"**
462	Faith Baptist Christian School
359	Jubilee Christian College
460	Capricornia School of Distance Education
462	Faith Baptist Christian School
464	Lutheran Ormeau Rivers District School
483	Australian Christian College
591 St Catherine's Catholic College The Whitsundays
625	Ambrose Treacy College

**Schools fixed manually**
23	Clairvaux Mackillop College
84	Emmanuel College
112	Marist College
113	Marist College			
115	St Aidan's Anglican Girls School
122 Moreton Bay Boys' College
184	St Joseph's College
251	OneSchool Global QLD
289	Aboriginal and Islander Independent Community School
320 Scots PGC College
322 St Joseph's School
341	Eagleby Learning Centre
384 Bayside Christian College
425	St Mary's Catholic College	
449	Marist College
477	The Cathedral College
534	The Cathedral School
572 Smithfield State High School
588	St Andrew's Catholic College
591 St Catherine's Catholic College
632 Cloncurry State School
678	Carlisle Adventist College

**Potential error**
656 Australian International School not found, with invalid postcode
657 Nauru Secondary School, with invalid postcode
664	Suzhou International Foreign Language School
665 Hangzhou Dong Fang School
667	Number 1 Middle School of Suzhou New District	
668	Qingdao Guokai Middle School
669	Anji Shangshu Private High School
  - Note: Most of these schools relate to overseas schools

## Duplicates 
* St Peter's Lutheran College - Springfield / St Peters Lutheran College Indooroopilly
* St Peter's Lutheran College = 4068
* St Peters Lutheran College - Springfield = 4300
* In `sep_all` both St Peters Lutheran College were combined into one school with wrong sep

```{r}
# Extract school's total no. of Senior Education Profile (SEP) for a given year 
sep_all <- school_outcomes %>% 
  select(completion_year,
         school_name = School,
         locality = Locality,
         sep = SEP)
```

```{r}
# Find duplicates in school outcomes data
# i.e. schools with > 1 locality
sep_duplicates <- sep_all %>% 
  # Extract all unique schools and locality
  distinct(school_name, locality) %>% 
  count(school_name) %>% 
  # Extract schools with >= 1 locality
  filter(n > 1) %>% 
  pull(school_name)

locality_no_dup <- sep_all %>% 
  filter(!school_name %in% sep_duplicates) %>% 
  select(-locality)

locality_dup <- sep_all %>% 
  filter(school_name %in% sep_duplicates) %>% 
  mutate(school_postcode = case_when(
    school_name == "Arcadia College" & locality == "Southport" ~ "4226", # School 
    school_name == "Arcadia College" & locality == "Varsity Lakes" ~ "4227", # PO BOX
    school_name == "Burnett Youth Learning Centre" & locality == "Bundaberg South" ~ "4670", # Bundaberg is in Elliott
    school_name == "Burnett Youth Learning Centre" & locality == "Elliott" ~ "4670",
    school_name == "Cairns School of Distance Education" & locality == "Manunda Cairns" ~ "4870", # school
    school_name == "Cairns School of Distance Education" & locality == "Cairns" ~ "4870", # PO BOX
    school_name == "Calvary Christian College" & locality == "Carbrook" ~ "4130",
    school_name == "Calvary Christian College" & locality == "Mount Louisa" ~ "4814",
    school_name == "Clontarf Beach State High School" & locality == "Clontarf" ~ "4019",
    school_name == "Clontarf Beach State High School" & locality == "Clontarf Beach" ~ "4019", # same address
    school_name == "Coolum State High School" & locality == "Coolum Beach" ~ "4573",
    school_name == "Coolum State High School" & locality == "Coolum" ~ "4573",
    school_name == "Craigslea State High School" & locality == "West Chermside" ~ "4032",
    school_name == "Craigslea State High School" & locality == "Chermside West" ~ "4032",
    school_name == "Emmaus College" & locality == "North Rockhampton" ~ "4701",
    school_name == "Emmaus College" & locality == "Jimboomba" ~ "4280",
    school_name == "Faith Lutheran College" & locality == "Plainland" ~ "4341",
    school_name == "Faith Lutheran College" & locality == "Victoria Point" ~ "4164",
    school_name == "Faith Lutheran College" & locality == "Thornlands" ~ "4164", # thornlands in vic point
    school_name == "Heatley Secondary College" & locality == "Heatley" ~ "4814",
    school_name == "Heatley Secondary College" & locality == "Townsville" ~ "4814", # heatley in townsville
    school_name == "Hymba Yumba Community Hub" & locality == "Springfield" ~ "4300",
    school_name == "Hymba Yumba Community Hub" & locality == "Springfield Central" ~ "4300", # same
    school_name == "Jabiru Community College" & locality == "Bracken Ridge" ~ "4017",
    school_name == "Jabiru Community College" & locality == "Zillmere" ~ "4034",
    school_name == "Marist College" & locality == "Ashgrove" ~ "4060",
    school_name == "Marist College" & locality == "Emerald" ~ "4720",
    school_name == "Marist College" & locality == "Paddington" ~ "4064", # permanently closed
    school_name == "Mt Maria College" & locality == "Mitchelton" ~ "4053",
    school_name == "Mt Maria College" & locality == "Petrie" ~ "4502",
    school_name == "North Rockhampton State High School" & locality == "Rockhampton North" ~ "4701", # rockhampton north is in frenchville
    school_name == "North Rockhampton State High School" & locality == "Frenchville" ~ "4701",
    school_name == "Pacific Lutheran College" & locality == "Meridan Plains" ~ "4551",
    school_name == "Pacific Lutheran College" & locality == "Birtinya" ~ "4551", # same
    school_name == "Pioneer State High School" & locality == "Andergrove Mackay" ~ "4740",
    school_name == "Pioneer State High School" & locality == "Andergrove" ~ "4740",
    school_name == "Saint Stephen's College" & locality == "Coomera" ~ "4209",
    school_name == "Saint Stephen's College" & locality == "Upper Coomera" ~ "4209",
    school_name == "Seton College" & locality == "Mount Gravatt East" ~ "4122",
    school_name == "Seton College" & locality == "East Mt Gravatt" ~ "4122",
    school_name == "Southern Cross Catholic College" & locality == "Scarborough" ~ "4020",
    school_name == "Southern Cross Catholic College" & locality == "Annandale" ~ "4814",
    school_name == "St Augustine's College" & locality == "Cairns" ~ "4870",
    school_name == "St Augustine's College" & locality == "Springfield" ~ "4300",
    school_name == "St Joseph's College" & locality == "Brisbane" ~ "4000",
    school_name == "St Joseph's College" & locality == "Toowoomba" ~ "4350",
    school_name == "St Joseph's College" & locality == "Stanthorpe" ~ "4380",
    school_name == "St Mary's College" & locality == "Toowoomba" ~ "4350",
    school_name == "St Mary's College" & locality == "Maryborough" ~ "4650",
    school_name == "St Peters Lutheran College" & locality == "Indooroopilly" ~ "4068",
    school_name == "St Peters Lutheran College" & locality == "Springfield Central" ~ "4300",
    school_name == "Thuringowa State High School" & locality == "Condon" ~ "4815",
    school_name == "Thuringowa State High School" & locality == "Townsville" ~ "4815",
    school_name == "Trinity Bay State High School" & locality == "Manunda" ~ "4870",
    school_name == "Trinity Bay State High School" & locality == "Manunda Cairns" ~ "4870",
    school_name == "Trinity College" & locality == "Beenleigh" ~ "4207",
    school_name == "Trinity College" & locality == "Gladstone" ~ "4680",
    school_name == "Western Suburbs State Special School" & locality == "Durack" ~ "4077",
    school_name == "Western Suburbs State Special School" & locality == "Inala" ~ "4077", # same
    school_name == "William Ross State High School" & locality == "Annandale" ~ "4814",
    school_name == "William Ross State High School" & locality == "Townsville" ~ "4814")) %>% # annandale is in city of townsville)
  select(-locality)
```

```{r}
sep_outcomes <- mathsci_all %>%
  # Convert school_postcode into character vector to facilitate relational join
  mutate(school_postcode = as.character(school_postcode)) %>% 
  left_join(locality_dup,
            by = c("completion_year",
                   "school_name",
                   "school_postcode")) %>% 
  left_join(locality_no_dup,
            by = c("completion_year",
                   "school_name")) %>% 
  # Remove overlaps due to relational join
  mutate(sep1 = case_when(
    is.na(sep.x) & is.na(sep.y) ~ "NA",
    is.na(sep.x)  ~ sep.y,
    is.na(sep.y)  ~ sep.x
  )) %>% 
  # Remove overlapping variables
  select(-c(sep.x, 
            sep.y)) 
```


```{r}
# Extract schools with two or more locations (postcodes)
enrolments_duplicates <- mathsci_all %>% 
  distinct(school_name, school_postcode) %>% 
  count(school_name) %>% 
  filter(n > 1) %>% 
  pull(school_name)

# Schools with only 1 location (postcode)
postcode_no_duplicates <- sep_all %>% 
  filter(!school_name %in% enrolments_duplicates) %>% 
  select(-locality)

# For duplicates, Manually include school_postcode to facilitate relational join
postcode_duplicates <- sep_all %>% 
  filter(school_name %in% enrolments_duplicates) %>% 
  mutate(school_postcode = case_when(
    school_name == "Australian International Islamic College" & locality == "Durack" ~ "4077",
    school_name == "Calvary Christian College" & locality == "Carbrook" ~ "4130",
    school_name == "Calvary Christian College" & locality == "Mount Louisa" ~ "4814",
    school_name == "Carinity Education" & locality == "Sunnybank" ~ "4109",
    school_name == "Emmaus College" & locality == "North Rockhampton" ~ "4701",
    school_name == "Emmaus College" & locality == "Jimboomba" ~ "4280",
    school_name == "Faith Lutheran College" & locality == "Plainland" ~ "4341",
    school_name == "Faith Lutheran College" & locality == "Victoria Point" ~ "4164",
    school_name == "Faith Lutheran College" & locality == "Thornlands" ~ "4164",
    school_name == "Grace Lutheran College" & locality == "Rothwell" ~ "4022",
    school_name == "Marist College" & locality == "Ashgrove" ~ "4060",
    school_name == "Marist College" & locality == "Emerald" ~ "4720",
    school_name == "Marist College" & locality == "Paddington" ~ "4000",
    school_name == "Mt Maria College" & locality == "Mitchelton" ~ "4053",
    school_name == "Mt Maria College" & locality == "Petrie" ~ "4502",
    school_name == "Southern Cross Catholic College" & locality == "Scarborough" ~ "4020",
    school_name == "Southern Cross Catholic College" & locality == "Annandale" ~ "4814",
    school_name == "St Augustine's College" & locality == "Cairns" ~ "4870",
    school_name == "St Augustine's College" & locality == "Springfield" ~ "4300",
    school_name == "St Joseph's College" & locality == "Brisbane" ~ "4000",
    school_name == "St Joseph's College" & locality == "Stanthorpe" ~ "4380",
    school_name == "St Joseph's College" & locality == "Toowoomba" ~ "4350",
    school_name == "St Mary's Catholic College" & locality == "Kingaroy" ~ "4610",
    school_name == "St Mary's College" & locality == "Toowoomba" ~ "4350",
    school_name == "St Mary's College" & locality == "Maryborough" ~ "4650",
    school_name == "St Patrick's College" & locality == "Gympie" ~ "4570",
    school_name == "St Ursula's College" & locality == "Yeppoon" ~ "4703",
    school_name == "Trinity College	" & locality == "Beenleigh" ~ "4207",
    school_name == "Trinity College	" & locality == "Gladstone" ~ "4680")) %>% 
  select(-locality)
```



```{r, warning=FALSE}
# Join enrolments data and SEP data
sep_outcomes <- sep_outcomes %>%
  # Convert school_postcode into character vector to facilitate relational join
  mutate(school_postcode = as.character(school_postcode)) %>% 
  # SEP data without duplicates
  left_join(postcode_no_duplicates, by = c("school_name",
                                      "completion_year")) %>%
  # SEP data with duplicates
  left_join(postcode_duplicates,
            by = c("completion_year",
                   "school_name",
                   "school_postcode")) %>% 
  # Remove overlaps due to relational join
  mutate(sep2 = case_when(
    is.na(sep.x) & is.na(sep.y) ~ "NA",
    is.na(sep.y)  ~ sep.x,
    is.na(sep.x)  ~ sep.y
  )) %>% 
  # Remove overlapping variables
  select(-c(sep.x, 
            sep.y)) %>% 
  # Remove overlapping variables from 
  mutate(sep = case_when(
    sep1 == "NA" ~ sep2,
    sep2 == "NA" ~ sep1,
    TRUE ~ sep1
  )) %>% 
  select(-c(sep1, sep2)) %>% 
  mutate(sep = as.numeric(sep))
```

```{r}
# Filter to years in which we have data for
sep_outcomes <- sep_outcomes %>% 
  filter(between(completion_year, 2008, 2019)) %>% 
  # Remove schools that are missing values for enrolments or SEP outcome (and vice versa)
  filter(!is.na(sep)) 

sep_outcomes %>% 
  filter(subject_name == "Chemistry") %>% 
  mutate(test = year_11_enrolments/sep) %>% 
  ggplot() +
  geom_jitter(aes(x = factor(completion_year),
                  y = test))
```













